---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

Before running the following code, it need to:
(1) install the three packages: ggplot2, dplyr, Hmisc
(2) set work space to a folder which contains this file and data file: "activity.csv"
## Loading and preprocessing the data
Firstly, we need to involve required packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(Hmisc)
```
Then read the CSV data
```{r}
activityData <- read.csv("activity.csv")
```
Then we preprocess the data
```{r}
cleanedData <- na.omit(activityData)
cleanedData$date <- as.Date(cleanedData$date)
```

## What is mean total number of steps taken per day?
Let's aggregate the data by a day
```{r}
stepsEachDay <- aggregate(steps ~ date, cleanedData, sum)
```
Then let's Plot statistic of steps

```{r echo=FALSE}
ggplot(stepsEachDay, aes(x = steps)) +
  geom_histogram(binwidth = 700, fill = "orange") +
  ggtitle("Steps Statistic") + 
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(x = "Steps", y = "Frequency")
```

The mean value of frequency of steps per day is 10766.19:
```{r echo=FALSE}
print(mean(stepsEachDay$steps))
```
The median value of frequency of steps per day is 10765:
```{r echo=FALSE}
print(median(stepsEachDay$steps))
```
## What is the average daily activity pattern?
Firstly, let's aggregate the data by interval
```{r}
stepsByInterval <- aggregate(steps ~ interval, cleanedData, mean)
```
plot time series plot of the average number of steps taken
```{r echo=FALSE}
ggplot(stepsByInterval, aes(x=interval, y=steps)) + 
  geom_line(color = "blue") +
  ggtitle("Time Series Plot of The Average Number of steps By Interval") + 
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(x = "Interval", y = "Average Number of steps")
```

The 5-minute interval that, on average, contains the maximum number of steps is 835:
```{r echo=FALSE}
print(stepsByInterval$interval[which.max(stepsByInterval$steps)])
```
## Imputing missing values
Calculate the number of NA
```{r echo=FALSE}
print(sum(is.na(activityData)))
```
use Package named: Hmisc to add mean of eliminating NA to fill the NA value
```{r}
prepareFilledNAData <- activityData
prepareFilledNAData$steps <- impute(activityData$steps, fun=mean)
```
aggregate the data by a day
```{r}
stepsEachDayWithFilledNA <- aggregate(steps ~ date, prepareFilledNAData, sum)
```

```{r echo=FALSE}
ggplot(stepsEachDayWithFilledNA, aes(x = steps)) +
  geom_histogram(binwidth = 700, fill = "orange") +
  ggtitle("Steps Statistic with filled NA by Mean") + 
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(x = "Steps", y = "Frequency")
```

After filling the NA, the mean value of frequency of steps per day: 10766.19
```{r echo=FALSE}
print(mean(stepsEachDay$steps))
```
After filling the NA, the median value of frequency of steps per day: 10765
```{r echo=FALSE}
print(median(stepsEachDay$steps))
```
## Are there differences in activity patterns between weekdays and weekends?
Preprocessing the date:
```{r}
prepareFilledNAData$date <- as.Date(prepareFilledNAData$date)
prepareFilledNAData$weekday <- weekdays(prepareFilledNAData$date)

prepareFilledNAData$dayType[prepareFilledNAData$weekday %in% c('Monday','Tuesday', 'Wednesday', 'Thursday', 'Friday')] <- "Weekday"
prepareFilledNAData$dayType[prepareFilledNAData$weekday %in% c('Saturday','Sunday')] <- "Weekend"

prepareFilledNAData$dayType <- as.factor(prepareFilledNAData$dayType)
```
aggregate by interval for weekday and weekend
```{r}
prepareFilledNADataByInterval <- aggregate(steps ~ interval + dayType, data=prepareFilledNAData, mean)
```
plot time series Of average steps By interval accross weekdays and weekends
```{r echo=FALSE}
ggplot(prepareFilledNADataByInterval , aes(x = interval , y = steps, color=dayType)) + 
  geom_line() + 
  facet_wrap(~dayType, ncol = 1, nrow=2) + 
  ggtitle("Average Number of steps By Interval Accross Weekdays & Weekends") + 
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(x = "Interval", y = "Average Number of steps")
```